Bayesian determination of the effect of a deep eutectic solvent on the structure of lipid monolayers

### 2018-11-06 reports/paper.tex
add ESI reference to figures

%
\begin{figure}[b]
\includegraphics[width=0.3\textwidth]{figures/head_groups}
\caption{\label{fig:heads}\small The two lipid classes with different head
groups compared in this study, where R indicates the hydrocarbon tail;
(a) phosphatidylglycerol (PG), (b) phosphocholine (PC).}
\end{figure}
%
We report on measurements of four phospholipids monolayers, namely
1,2-dipalmitoyl-sn-glycero-3-phosphocholine (DPPC, C$_{16}$ tails),
1,2-dimyristoyl-sn-glycero-3-phosphocholine (DMPC, C$_{14}$ tails),
1,2-dilauroyl-sn-glycero-3-phosphocholine (DLPC, C$_{12}$ tails) and
1,2-dimyristoyl-sn-glycero-3-phospho-(1'-rac-glycerol)
(DMPG, C$_{14}$ tails), at the air-DES (1:2 choline choride:glycerol)
interface.
In contrast to many previous studies\cite{Mohwald1990,Kewalramani2010,
Bayerl1990,Johnson1991,Clifton2012,Helm1987,Daillant1990}, we have developed
a chemically-consistent model that allows for the
co-refinement of reflectometry measurements at different surface pressure
and makes no assumption of the volume of the lipid head, $V_h$, or tail, $V_t$ \cite{SI2018}.
Instead these parameters were allowed to vary for each lipid while being
constrained to be self-consistent over different surface pressures in the
same phase; Liquid-Condensed (LC) for DPPC and Liquid-Expanded (LE) for
DMPC, DMPG and DLPC.
This model was required because we cannot assume that the electrostatic
interactions between lipid head and the solvent are the same in the DES and
in water. This may therefore influence the effective head volume which means
we cannot rely on the literature values derived from measurements
on or in water \cite{SI2018}.
Furthermore, it is known that, on water, increased surface pressure and
associated LE-LC phase transitions lead to a compression of the lipid tail
volume\cite{Marsh2010,Small1984} and this compaction has not necessarily
been accounted for in the literature\cite{Campbell2018}.
Our approach avoids this issue by making no assumption about the molecular
volumes and only considering surface pressures that we believe to be in the
same phase.

%
\begin{figure}
	\centering
  \includegraphics[width=0.45\textwidth]{figures/dlpc_ref_sld}
	\includegraphics[width=0.45\textwidth]{figures/dmpc_ref_sld}
	\includegraphics[width=0.45\textwidth]{figures/dppc_ref_sld}
	\includegraphics[width=0.45\textwidth]{figures/dmpg_ref_sld}
	\caption{\small The XRR profiles (left) and SLD profiles (right) for each
  of the four lipids; (a) DLPC, (b) DMPC, (c) DPPC, (d) DMPG, at the four
  measured surface pressures; see legend above each plot. The different
  surface pressure XRR profiles have been offset in the $y$-axis by an order
  of magnitude and SLD profiles offset in the $y$-axis by
  \SI{5e-6}{\per\square\angstrom}, for clarity.}
	\label{fig:lipids}
\end{figure}
%
Our model has more variables than we can uniquely fit with limited data.
In reflectometry the usual approach to this problem is to collect more
equivalent dataset with varying deuteration (NR). For reasons of cost and
practicality this was not possible in this work.
The other approach is to constrain the model so that the number of fitted
variables are reduced, and is the approach taken here.
To do this we add the constraint that a single tail volume can be used for
all surface pressures. To justify this we need to be sure that the lipids
remain in the same phase. On water this is can be demonstrated with a
Langmuir isotherm.
However, while we have confidence that the individual surface pressures
measured were reliable, we were unable to collect consistent Langmuir
isotherm measurements, due to the high viscosity of the DES.
This inconsistency means that we cannot be confident in the phase of the
lipids based on the surface pressure alone.
Instead we have used grazing incidence X-ray diffraction to confirm the
phases of DMPC and DPPC at \SI{30}{\milli\newton\per\meter}.
DPPC was found to be in the LC phase and DMPC in the LE phase at room
temperature for the surface pressures measured (see Section \ref{sec:gixd}).
We assume that DMPG and DLPC are also in the LE phase since there is no
reason to believe that the phase behaviour in these systems differs
significantly from DMPC at the same temperature.

@@

Figure \ref{fig:lipresults} shows the tail layer thickness variation with
surface pressure. For DPPC and DMPC a plateau in thickness is reached at
\SI{25}{\milli\newton\per\meter} and \SI{30}{\milli\newton\per\meter}
respectively. Presumably a similar plateau would be seen for DMPG and DLPC
at higher pressures.
This phenomenon has been noted before for DMPC\cite{Bayerl1990} and
DPPC\cite{Campbell2018} at the air-water interface.
%
\begin{figure}
	\centering
  \includegraphics[width=0.45\textwidth]{figures/dlpc_vh_dt_phi}
	\includegraphics[width=0.45\textwidth]{figures/dmpc_vh_dt_phi}
	\includegraphics[width=0.45\textwidth]{figures/dppc_vh_dt_phi}
	\includegraphics[width=0.45\textwidth]{figures/dmpg_vh_dt_phi}
	\caption{\small The PDFs of the head volume (left) and variation of
  $d_t$ (squares) and $\phi_h$ (circles) with surface pressure for each of
  the four lipids; (a) DLPC, (b) DMPC, (c) DPPC, (d) DMPG. The values of
  $d_t$ were found from the appropriate values of $\theta_t$ using
  Eqn. \ref{equ:tl}.}
	\label{fig:lipresults}
\end{figure}
%

@@

The volumes determined for the head and tails, and the thickness of the head
layers, for DMPC and DPPC were used to constrain the fitting of corresponding
NR data (Figure \ref{fig:neutron}). This leaves only two variables,
$\theta_t$ and $\sigma$ to fit this data.
The agreement is good and a confirmation that the volumes derived from XRR
are consistent with the NR data.
Furthermore, the observed trends with surface pressure are also consistent
which adds some confidence in our conclusions.
%
\begin{figure}
	\centering
  \includegraphics[width=0.45\textwidth]{figures/dmpc_20n_ref_sld}
	\includegraphics[width=0.45\textwidth]{figures/dppc_20n_ref_sld}
	\caption{\small The NR and SLD profiles at a surface pressure of
  20 mNm$^{-1}$ for two contrasts (see legend above each plot); (a) DMPC,
  (b) DPPC. The NR profiles have been offset in the $y$-axis by an order of
  magnitude and SLD profiles offset in the $y$-axis by
  \SI{5e-6}{\per\square\angstrom}, for clarity.}
	\label{fig:neutron}
\end{figure}
%
%
\begin{table}
  \caption{\label{tab:neutron} The best-fit values, and associated 95 \%
  confidence intervals for the varying parameters in the co-refined NR models.
  The values of $d_t$ were found from the appropriate values of $\theta_t$
  using Eqn. \ref{equ:tl}, and the values of $\phi_h$ were found using Eqn.
  \ref{equ:phih}.}
  \begin{ruledtabular}
	\begin{tabular}{ccccc}
    Lipid & \multicolumn{2}{c}{d$_{54}$-DMPC} & \multicolumn{2}{c}{d$_{62}$-DPPC} \\
    SP/mNm$^{-1}$ & 20 & 25 & 15 & 20 \\
    \hline
    $\theta_t$/$^\circ$ & \input{../output/dmpc/angle20_neutron.txt} &
    \input{../output/dmpc/angle25_neutron.txt} &
    \input{../output/dppc/angle15_neutron.txt} &
    \input{../output/dppc/angle20_neutron.txt} \\
    $\sigma_{t,h,s}$/\AA & \input{../output/dmpc/rough20_neutron.txt} &
    \input{../output/dmpc/rough25_neutron.txt} &
    \input{../output/dppc/rough15_neutron.txt} &
    \input{../output/dppc/rough20_neutron.txt} \\
    \hline
    $\phi_h$/$\times10^{-2}$ & \input{../output/dmpc/solh20_neutron.txt} &
    \input{../output/dmpc/solh25_neutron.txt} &
    \input{../output/dppc/solh15_neutron.txt} &
    \input{../output/dppc/solh20_neutron.txt} \\
    $d_t$/\AA & \input{../output/dmpc/tail20_neutron.txt} &
    \input{../output/dmpc/tail25_neutron.txt} &
    \input{../output/dppc/tail15_neutron.txt} &
    \input{../output/dppc/tail20_neutron.txt} \\
	\end{tabular}
  \end{ruledtabular}
\end{table}
%

>>>>

%
\begin{figure}[b]
\includegraphics[width=0.3\textwidth]{figures/head_groups}
\caption{\label{fig:heads}\small The two lipid classes with different head
groups compared in this study, where R indicates the hydrocarbon tail;
(a) phosphatidylglycerol (PG), (b) phosphocholine (PC) \cite{SI2018}.}
\end{figure}
%
We report on measurements of four phospholipids monolayers, namely
1,2-dipalmitoyl-sn-glycero-3-phosphocholine (DPPC, C$_{16}$ tails),
1,2-dimyristoyl-sn-glycero-3-phosphocholine (DMPC, C$_{14}$ tails),
1,2-dilauroyl-sn-glycero-3-phosphocholine (DLPC, C$_{12}$ tails) and
1,2-dimyristoyl-sn-glycero-3-phospho-(1'-rac-glycerol)
(DMPG, C$_{14}$ tails), at the air-DES (1:2 choline choride:glycerol)
interface.
In contrast to many previous studies\cite{Mohwald1990,Kewalramani2010,
Bayerl1990,Johnson1991,Clifton2012,Helm1987,Daillant1990}, we have developed
a chemically-consistent model that allows for the
co-refinement of reflectometry measurements at different surface pressure
and makes no assumption of the volume of the lipid head, $V_h$, or tail, $V_t$ \cite{SI2018}.
Instead these parameters were allowed to vary for each lipid while being
constrained to be self-consistent over different surface pressures in the
same phase; Liquid-Condensed (LC) for DPPC and Liquid-Expanded (LE) for
DMPC, DMPG and DLPC.
This model was required because we cannot assume that the electrostatic
interactions between lipid head and the solvent are the same in the DES and
in water. This may therefore influence the effective head volume which means
we cannot rely on the literature values derived from measurements
on or in water \cite{SI2018}.
Furthermore, it is known that, on water, increased surface pressure and
associated LE-LC phase transitions lead to a compression of the lipid tail
volume\cite{Marsh2010,Small1984} and this compaction has not necessarily
been accounted for in the literature\cite{Campbell2018}.
Our approach avoids this issue by making no assumption about the molecular
volumes and only considering surface pressures that we believe to be in the
same phase.

%
\begin{figure}
	\centering
  \includegraphics[width=0.45\textwidth]{figures/dlpc_ref_sld}
	\includegraphics[width=0.45\textwidth]{figures/dmpc_ref_sld}
	\includegraphics[width=0.45\textwidth]{figures/dppc_ref_sld}
	\includegraphics[width=0.45\textwidth]{figures/dmpg_ref_sld}
	\caption{\small The XRR profiles (left) and SLD profiles (right) for each
  of the four lipids; (a) DLPC, (b) DMPC, (c) DPPC, (d) DMPG, at the four
  measured surface pressures; see legend above each plot. The different
  surface pressure XRR profiles have been offset in the $y$-axis by an order
  of magnitude and SLD profiles offset in the $y$-axis by
  \SI{5e-6}{\per\square\angstrom}, for clarity \cite{SI2018}.}
	\label{fig:lipids}
\end{figure}
%
Our model has more variables than we can uniquely fit with limited data.
In reflectometry the usual approach to this problem is to collect more
equivalent dataset with varying deuteration (NR). For reasons of cost and
practicality this was not possible in this work.
The other approach is to constrain the model so that the number of fitted
variables are reduced, and is the approach taken here.
To do this we add the constraint that a single tail volume can be used for
all surface pressures. To justify this we need to be sure that the lipids
remain in the same phase. On water this is can be demonstrated with a
Langmuir isotherm.
However, while we have confidence that the individual surface pressures
measured were reliable, we were unable to collect consistent Langmuir
isotherm measurements, due to the high viscosity of the DES.
This inconsistency means that we cannot be confident in the phase of the
lipids based on the surface pressure alone.
Instead we have used grazing incidence X-ray diffraction to confirm the
phases of DMPC and DPPC at \SI{30}{\milli\newton\per\meter}.
DPPC was found to be in the LC phase and DMPC in the LE phase at room
temperature for the surface pressures measured (see Section \ref{sec:gixd}).
We assume that DMPG and DLPC are also in the LE phase since there is no
reason to believe that the phase behaviour in these systems differs
significantly from DMPC at the same temperature.

@@

Figure \ref{fig:lipresults} shows the tail layer thickness variation with
surface pressure. For DPPC and DMPC a plateau in thickness is reached at
\SI{25}{\milli\newton\per\meter} and \SI{30}{\milli\newton\per\meter}
respectively. Presumably a similar plateau would be seen for DMPG and DLPC
at higher pressures.
This phenomenon has been noted before for DMPC\cite{Bayerl1990} and
DPPC\cite{Campbell2018} at the air-water interface.
%
\begin{figure}
	\centering
  \includegraphics[width=0.45\textwidth]{figures/dlpc_vh_dt_phi}
	\includegraphics[width=0.45\textwidth]{figures/dmpc_vh_dt_phi}
	\includegraphics[width=0.45\textwidth]{figures/dppc_vh_dt_phi}
	\includegraphics[width=0.45\textwidth]{figures/dmpg_vh_dt_phi}
	\caption{\small The PDFs of the head volume (left) and variation of
  $d_t$ (squares) and $\phi_h$ (circles) with surface pressure for each of
  the four lipids; (a) DLPC, (b) DMPC, (c) DPPC, (d) DMPG. The values of
  $d_t$ were found from the appropriate values of $\theta_t$ using
  Eqn. \ref{equ:tl} \cite{SI2018}.}
	\label{fig:lipresults}
\end{figure}
%

@@

The volumes determined for the head and tails, and the thickness of the head
layers, for DMPC and DPPC were used to constrain the fitting of corresponding
NR data (Figure \ref{fig:neutron}). This leaves only two variables,
$\theta_t$ and $\sigma$ to fit this data.
The agreement is good and a confirmation that the volumes derived from XRR
are consistent with the NR data.
Furthermore, the observed trends with surface pressure are also consistent
which adds some confidence in our conclusions.
%
\begin{figure}
	\centering
  \includegraphics[width=0.45\textwidth]{figures/dmpc_20n_ref_sld}
	\includegraphics[width=0.45\textwidth]{figures/dppc_20n_ref_sld}
	\caption{\small The NR and SLD profiles at a surface pressure of
  20 mNm$^{-1}$ for two contrasts (see legend above each plot); (a) DMPC,
  (b) DPPC. The NR profiles have been offset in the $y$-axis by an order of
  magnitude and SLD profiles offset in the $y$-axis by
  \SI{5e-6}{\per\square\angstrom}, for clarity \cite{SI2018}.}
	\label{fig:neutron}
\end{figure}
%
%
\begin{table}
  \caption{\label{tab:neutron} The best-fit values, and associated 95 \%
  confidence intervals for the varying parameters in the co-refined NR models.
  The values of $d_t$ were found from the appropriate values of $\theta_t$
  using Eqn. \ref{equ:tl}, and the values of $\phi_h$ were found using Eqn.
  \ref{equ:phih}.}
  \begin{ruledtabular}
	\begin{tabular}{ccccc}
    Lipid & \multicolumn{2}{c}{d$_{54}$-DMPC} & \multicolumn{2}{c}{d$_{62}$-DPPC} \\
    SP/mNm$^{-1}$ & 20 & 25 & 15 & 20 \\
    \hline
    $\theta_t$/$^\circ$ & \input{../output/dmpc/angle20_neutron.txt} &
    \input{../output/dmpc/angle25_neutron.txt} &
    \input{../output/dppc/angle15_neutron.txt} &
    \input{../output/dppc/angle20_neutron.txt} \\
    $\sigma_{t,h,s}$/\AA & \input{../output/dmpc/rough20_neutron.txt} &
    \input{../output/dmpc/rough25_neutron.txt} &
    \input{../output/dppc/rough15_neutron.txt} &
    \input{../output/dppc/rough20_neutron.txt} \\
    \hline
    $\phi_h$/$\times10^{-2}$ & \input{../output/dmpc/solh20_neutron.txt} &
    \input{../output/dmpc/solh25_neutron.txt} &
    \input{../output/dppc/solh15_neutron.txt} &
    \input{../output/dppc/solh20_neutron.txt} \\
    $d_t$/\AA & \input{../output/dmpc/tail20_neutron.txt} &
    \input{../output/dmpc/tail25_neutron.txt} &
    \input{../output/dppc/tail15_neutron.txt} &
    \input{../output/dppc/tail20_neutron.txt} \\
	\end{tabular}
  \end{ruledtabular}
\end{table}
%
